---
title: "Final_Project"
author: "Makoto Fukumoto"
date: "12/11/2016"
output: pdf_document
---

# Import 2012 and 2016 election data that Tony McGovern has compiled by scraping through townhall.com home page. (Official result is not yet available in county level as of December 11, 2016)

```{r}
# import necessary package
library(RCurl)

# Get URL 
x <- getURL("https://raw.githubusercontent.com/tonmcg/County_Level_Election_Results_12-16/master/US_County_Level_Presidential_Results_12-16.csv")

# Read it in CSV
y <- read.csv(text = x)

# See the summary
summary(y)
```

# Create a couple of variables to use in the analysis

```{r}
# The growth in the vote count 
y$puregrowth <- y$votes_gop_2016 - y$votes_gop_2012

# The growth in the voting share
y$pergrowth <- y$per_gop_2016 - y$per_gop_2012

# THe growth in the winning (losing) margin
y$margingrowth <- -(y$per_point_diff_2016 - y$per_point_diff_2012)

# Visualize the data
hist(y$margingrowth)
summary(y$margingrowth)

# ALaska was not available yet, so there are 29 NAs.

# Write the arranged data into CSV file
write.csv(y, file = "election.csv")

```

# Read and clean the Census data

```{r}
# Read the census data into R
census <- read.csv('~/Desktop/PS239T/07_APIs/census2014.csv')

# See summary
summary(census)

# Create some bew variables 
census$homegrown <- census$BornInState/census$Population
census$same_house <- census$SameHouse/census$Population

  # Recent move in and move out; aggreagete them and devide by population
census$move_in<- (census$FromOtherState+census$FromOtherCounty) /census$Population
census$move_out <- (census$ToOtherState+census$ToOtherCounty) /census$Population

  # The following variables had NAs and thus it was easier to merge in R rather than in Python
census$youth <- census$X18to24/census$Population
census$elder <- census$X65to74/census$Population

```

# Do the same for 2010 data

```{r}


census2010 <- read.csv('~/Desktop/PS239T/07_APIs/census2010.csv')

summary(census2010)

census2010$homegrown10 <- census2010$BornInState10/census2010$Population10
census2010$same_house10 <- census2010$SameHouse/census2010$Population10
census2010$move_in10<- (census2010$FromOtherState10+census2010$FromOtherCounty10) /census2010$Population10
census2010$move_out10 <- (census2010$ToOtherState10+census2010$ToOtherCounty10) /census2010$Population10
census2010$youth10 <- census2010$X18to24_10/census2010$Population10
census2010$elder10 <- census2010$X65to74_10/census2010$Population10

```

# Merge the files with County name and State FIPS code (tried to merge wiwhtcounty name and found many counties with the same name, so merged by two variavles instead)

```{r}
total<- merge(census,census2010,by=c("county_name","state_fips"))
total2 <- merge(y,total,by=c("county_name","state_fips"))
```

# Create Population growth data and Income growth data, using the merged data

```{r}

total2$pop_growth <- (total2$Population - total2$Population10)/total2$Population10
total2$income_growth <- (total2$MedianIncome - total2$MedianIncome10)/total2$MedianIncome10

summary(total2)
```

# Retrieve the data for 8 swing states

```{r}

virginia2 <- total2[which(total2$State=='Virginia'),]
michigan2 <- total2[which(total2$State=='Michigan'),]
penn2 <- total2[which(total2$State=='Pennsylvania'),]
florida2 <- total2[which(total2$State=='Florida'),]
wisconsin2 <- total2[which(total2$State=='Wisconsin'),]
ohio2 <- total2[which(total2$State=='Ohio'),]
nc2 <- total2[which(total2$State=='North Carolina'),]
nh2 <- total2[which(total2$State=='New Hampshire'),]
swing2 <- rbind(virginia2,michigan2,penn2,florida2,wisconsin2,ohio2,nc2,nh2)

```

# Visualization 

```{r}

# Import the packages necessary for visualization 

require(ggplot2)
require(gridExtra)

```

# Covariates and Trump margin - Romney margin; 

```{r}

# Hispanic * Trump

ggplot(swing2, aes(x = hisp_r, y = margingrowth))  + 
  ggtitle("Hispanic and Trump margin - Romney margin") + 
  geom_point(aes(colour = white_r) )+ 
  scale_color_continuous(name="White", low = 'pink', high = "red")  

# Just for Florida

ggplot(florida2, aes(x = hisp_r, y = margingrowth))  + 
  ggtitle("Florida: Hispanic and Trump margin - Romney margin") + 
  geom_point(aes(colour = white_r) )+ 
  scale_color_continuous(name="White", low = 'pink', high = "red")  

# Median Income * Trump 

ggplot(swing2, aes(x = MedianIncome, y = margingrowth))  + 
  ggtitle("Median Income and Trump margin - Romney margin") + 
  geom_point(aes(colour = Management)) + 
  scale_color_continuous(name="Management Occupation", low = 'pink', high = "red")  

# Just for Virginia
ggplot(virginia2, aes(x = MedianIncome, y = margingrowth))  + 
  ggtitle("Virginia: Median Income and Trump margin - Romney margin") + 
  geom_point(aes(colour = Management)) + 
  scale_color_continuous(name="Management Occupation", low = 'pink', high = "red")  

# Education * Trump
ggplot(swing2, aes(x = univ_r, y = margingrowth))  + 
  ggtitle("Bachelor degree holder's share and Trump's margin compared to Romney's") + 
  geom_point(aes(colour = MedianIncome)) + 
  scale_color_continuous(name="High School", low = 'pink', high = "red") 

# Just for North Carolina
ggplot(nc2, aes(x = univ_r, y = margingrowth))  + 
  ggtitle("NC: Bachelor degree holder's share and Trump's margin compared to Romney's") + 
  geom_point(aes(colour = MedianIncome)) + 
  scale_color_continuous(name="High School", low = 'pink', high = "red") 

# Elderly * Trump 

ggplot(swing2, aes(x = elder, y = margingrowth))  + 
  ggtitle("Elderly and Trump's margin compared to Romney's") + 
  geom_point(aes(colour = youth)) + 
  scale_color_continuous(name="youth", low = 'pink', high = "red")  

# Just for New Hampshire

ggplot(nh2, aes(x = elder, y = margingrowth))  + 
  ggtitle("Elderly and Trump's margin compared to Romney's") + 
  geom_point(aes(colour = youth)) + 
  scale_color_continuous(name="youth", low = 'pink', high = "red")  


# Manufacture * Trump

ggplot(swing2, aes(x = manufacture_r, y = margingrowth))  + 
  ggtitle("Manufacture and Trump's margin compared to Romney's") + 
  geom_point(aes(colour = agriculture_r)) + 
  scale_color_continuous(name="Agriculture", low = 'pink', high = "red")  

# Just for Michigan

ggplot(michigan2, aes(x = manufacture_r, y = margingrowth))  + 
  ggtitle("Michigan: Manufacture and Trump's margin compared to Romney's") + 
  geom_point(aes(colour = agriculture_r)) + 
  scale_color_continuous(name="Agriculture", low = 'pink', high = "red")  

ls(total2)

# Unemployment rate * Trump

ggplot(swing2, aes(x = unemp_r, y = margingrowth))  + 
  ggtitle("Unemployment rate and Trump's margin compared to Romney's") + 
  geom_point(aes(colour = active_r)) + 
  scale_color_continuous(name="Activity rate", low = 'pink', high = "red")  

# Just for Ohio

ggplot(ohio2, aes(x = unemp_r, y = margingrowth))  + 
  ggtitle("Ohio: Unemployment rate and Trump's margin compared to Romney's") + 
  geom_point(aes(colour = active_r)) + 
  scale_color_continuous(name="Activity rate", low = 'pink', high = "red")  

```

# Main point: Population growth and increasing margin

```{r}

# Population growth * Trump vote
ggplot(swing2, aes(x = pop_growth, y = margingrowth))  + 
  ggtitle("Population growth and Trump's margin compared to Romney's") + 
  geom_point(aes(colour = move_in)) + 
  scale_color_continuous(name="Recent move in", low = 'pink', high = "red")  +
  geom_smooth(method = "lm", se = FALSE)

# Just for Virginia

ggplot(virginia2, aes(x = pop_growth, y = margingrowth))  + 
  ggtitle("Virginia: Population growth and Trump's margin compared to Romney's") + 
  geom_point(aes(colour = move_in)) + 
  scale_color_continuous(name="Recent move in", low = 'pink', high = "red")   +
  geom_smooth(method = "lm", se = FALSE)

# Income growth * Trump vote

ggplot(swing2, aes(x = income_growth, y = margingrowth))  + 
  ggtitle("Median Income growth and Trump's margin compared to Romney's") + 
  geom_point(aes(colour = active_r)) + 
  scale_color_continuous(name="Activity rate", low = 'pink', high = "red")

# Just for Virginia 

ggplot(virginia2, aes(x = income_growth, y = margingrowth))  + 
  ggtitle("Virginia: Median Income growth and Trump's margin compared to Romney's") + 
  geom_point(aes(colour = active_r)) + 
  scale_color_continuous(name="Activity rate", low = 'pink', high = "red")  


```

# Use dependent variabales other than margins

```{r}

# Voting share 
ggplot(swing2, aes(x = pop_growth, y = pergrowth))  + 
  ggtitle("Population growth and Trump's share compared to Romney's") + 
  geom_point(aes(colour = move_in)) + 
  scale_color_continuous(name="Recent move in", low = 'pink', high = "red")  +
  geom_smooth(method = "lm", se = FALSE)

# Vote count 
ggplot(swing2, aes(x = pop_growth, y = margingrowth))  + 
  ggtitle("Population growth and Trump's margin compared to Romney's") + 
  geom_point(aes(colour = move_in)) + 
  scale_color_continuous(name="Recent move in", low = 'pink', high = "red")  +
  geom_smooth(method = "lm", se = FALSE)

```

# Go deeper than "population decrease" 

```{r}
ggplot(swing2, aes(x = homegrown, y = margingrowth))  + 
  ggtitle("Population born in the state and Trump's margin compared to Romney's") + 
  geom_point(aes(colour = move_in)) + 
  scale_color_continuous(name="Recent move in", low = 'pink', high = "red")  +
  geom_smooth(method = "lm", se = FALSE)

ggplot(swing2, aes(x = same_house, y = margingrowth))  + 
  ggtitle("Population living in the Same house 1 year ago and Trump's margin compared to Romney's") + 
  geom_point(aes(colour = move_out)) + 
  scale_color_continuous(name="Recent move out", low = 'pink', high = "red")  

ggplot(swing2, aes(x = move_in, y = margingrowth))  + 
  ggtitle("Population recently moved in and Trump's margin compared to Romney's") + 
  geom_point(aes(colour = move_out)) + 
  scale_color_continuous(name="Recent move out", low = 'pink', high = "red")  

```


# Multivaraiate regression

```{r}
library(stargazer)

# Full model 

lm0 <- lm(margingrowth~pop_growth+income_growth+homegrown+same_house+move_in+move_out+
            univ_r+highschool_r+white_r+hisp_r+black_r+unemp_r+active_r+elder+youth+manufacture_r+agriculture_r+management_r+MedianIncome+MedianAge+homeowner_r+Foodstamp,
          data=swing2)
summary(lm0)

# Omit some variables that are closely related to each other 

lm1 <- lm(margingrowth~pop_growth+income_growth+
            univ_r+hisp_r+black_r+unemp_r+active_r+elder+youth+manufacture_r+agriculture_r+MedianIncome+MedianAge+homeowner_r+Foodstamp,
          data=swing2)
summary(lm1)

lm2 <- lm(margingrowth~pop_growth+income_growth+homegrown+move_in+
            univ_r+hisp_r+black_r+active_r+elder+manufacture_r+agriculture_r+MedianIncome+MedianAge+homeowner_r+Foodstamp,
          data=swing2)
summary(lm2)

# USA as a whole, not Swing states 

lm3 <- lm(margingrowth~pop_growth+income_growth+
            univ_r+hisp_r+black_r+unemp_r+active_r+elder+youth+manufacture_r+agriculture_r+MedianIncome+MedianAge+homeowner_r+Foodstamp,
          data=total2)
summary(lm3)


stargazer(lm0, lm1, lm2, lm3, type = "text", title="Regression Results")
```

